Basics of cubic spline models

### 2021-11-24 text/21/cubic-splines.Rmd
Converted plot to ggplot in cubic-splines.Rmd

```{r}
x <- c(0:20, setdiff(seq(0, 20, by=1/64), 0:20))
xm <- data.frame(
  c1 =rep(1, length(x)),
  c2 =x,
  c3 =x^2,
  c4 =x^3,
  c5 =(x> 5)*rep(1, length(x)),
  c6 =(x> 5)*(x- 5),
  c7 =(x> 5)*(x- 5)^2,
  c8 =(x> 5)*(x- 5)^3,
  c9 =(x>10)*rep(1, length(x)),
  c10=(x>10)*(x-10),
  c11=(x>10)*(x-10)^2,
  c12=(x>10)*(x-10)^3,
  c13 =(x>15)*rep(1, length(x)),
  c14=(x>15)*(x-15),
  c15=(x>15)*(x-15)^2,
  c16=(x>15)*(x-15)^3,
  y=NA
)
xm$y[1:21] <- y
```

```{r}
kable(xm[1:21, 1:4])
single_cubic_model <- lm(y~c2+c3+c4, data=xm)
single_cubic_model %>% 
  augment(newdata=xm) %>%
  ggplot(aes(x, y)) + 
    geom_point() +
    geom_line(aes(y=.fitted)) +
    ggtitle("Single cubic fit to the data")
```

>>>>

```{r}
x <- c(0:20, setdiff(seq(0, 20, by=1/64), 0:20))
```

You can get a variety of splines by defining constant, linear, quadratic, and cubic terms and then shift those functions to the right with zeros to the left.

```{r}
xm <- data.frame(
  c1 =rep(1, length(x)),
  c2 =x,
  c3 =x^2,
  c4 =x^3,
  c5 =(x> 5)*rep(1, length(x)),
  c6 =(x> 5)*(x- 5),
  c7 =(x> 5)*(x- 5)^2,
  c8 =(x> 5)*(x- 5)^3,
  c9 =(x>10)*rep(1, length(x)),
  c10=(x>10)*(x-10),
  c11=(x>10)*(x-10)^2,
  c12=(x>10)*(x-10)^3,
  c13 =(x>15)*rep(1, length(x)),
  c14=(x>15)*(x-15),
  c15=(x>15)*(x-15)^2,
  c16=(x>15)*(x-15)^3,
  y=NA
)
xm$y[1:21] <- y
```

```{r}
lb1 <- rep(c("intercept", "linear term", "quadratic term", "cubic term"), 4)
lb2 <- rep(c("for the full range", "restarted at x=5", "restarted at x=10", "restarted at x=15"), each=4)
for (j in 1:16) {
  xm[1:21, ] %>%
    ggplot(aes(c2, .data[[paste0("c", j)]])) +
      ggtitle(paste(lb1[j], lb2[j])) +
      geom_point() +
      xlab(" ") +
      ylab(" ") -> cplot
  plot(cplot)
}
```

### Single cubic polynomial

This is the matrix of independent variables for a single cubic polynomial.

```{r}
kable(xm[1:21, 1:4])
```

The graph shown below represents the best fitting single cubic polynomial.

```{r}
single_cubic_model <- lm(y~c2+c3+c4, data=xm)
single_cubic_model %>% 
  augment(newdata=xm) %>%
  ggplot(aes(x, y)) + 
    geom_point() +
    geom_line(aes(y=.fitted)) +
    ggtitle("Single cubic fit to the data")
```

### Separate cubic polynomials
